A preliminary constraint grammar for Russian

### 2015-05-04 2015-nodalida-ruscg/ruscg.tex
revising
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60217 72bbbca6-d526-0410-a7d9-f06f51895060

This paper presents a preliminary constraint grammar for Russian. The main 
objective of the constraint grammar is to produce a high recall grammar to serve
as input into other natural language processing tasks.
\rr{R3: What is the rationale for choosing high recall at the expense of low precision?
Was it a deliberate choice or a consequence of limited development time? In my
experience, this strategy is good for corpus queries in linguistic research
(fewer relevant constructions are missed) but might be problematic for language
technology applications, which normally require full disambiguation. Since it is
stated in the introduction that the principal objective of the grammar is to
produce input to other language processing tasks, this question is especially
relevant and warrants some discussion.}

@@

\needspace{6\baselineskip} % keeps example all together (no orphan line) 
\ex. Morphosyntactically congruent homographs \label{ex:MScongruent}
\a. \rus{з\'{a}мок} \emph{z\'{a}mok} `castle.\textsubscript{SG-NOM}'\\
	\rus{зам\'{о}к} \emph{zam\'{o}k} `lock.\textsubscript{SG-NOM}'
\b. \rus{з\'{a}мков} \emph{z\'{a}mkov} `castle.\textsubscript{PL-GEN}'\\
	\rus{замк\'{о}в} \emph{zamk\'{o}v} `lock.\textsubscript{PL-GEN}'
\c. etc. \ldots\\ \rr{is this clear that the paradigm continues, but we don't show it?}

@@

\subsection{Corpus}
In order to evaluate the grammar we hand-annotated 10,150 words of Russian text
from Wikipedia articles, public domain literature and freely-available news sources. The
annotated texts are available online under the {\sc cc-by-sa} licence.\footnote{\url{https://svn.code.sf.net/p/apertium/svn/languages/apertium-rus/texts/}}
\rr{is this a good url? Do we need to put a COPYING file in that folder?}

@@

\subsection{Intrinsic}
%FRAN
\rr{R1 does not understand "intrinsic/extrinsic" (and neither do I) Should we rename them to be qualitative/quantitative?}
\rr{R1: Add some text introducing intrinsic evaluation. What is this section doing?}

@@

\subsection{Extrinsic}
% ROB accent placement task

>>>>

This paper presents a preliminary constraint grammar for Russian. The main 
objective of the constraint grammar is to produce a high recall grammar to serve
as input into other natural language processing tasks. There are two reasons to
maintain high recall. First, one of the primary applications for this constraint grammar
is computer-assisted language learning. In the domain, erroneous analyses can lead to
significant frustration for learners. Therefore, it is important to limit disambiguation
to cases that can be resolved with high confidence.
Second, it is frequently the case that competing readings can
be distinguished only by considering idiosyncratic collocational information. For such cases, 
we expect that probabilistic approaches are both more effective and simpler to implement.

@@

\needspace{6\baselineskip} % keeps example all together (no orphan line) 
\ex. Morphosyntactically congruent homographs \label{ex:MScongruent}
\a. \rus{з\'{a}мок} \emph{z\'{a}mok} `castle.\textsubscript{SG-NOM}'\\
	\rus{зам\'{о}к} \emph{zam\'{o}k} `lock.\textsubscript{SG-NOM}'
\b. \rus{з\'{a}мков} \emph{z\'{a}mkov} `castle.\textsubscript{PL-GEN}'\\
	\rus{замк\'{о}в} \emph{zamk\'{o}v} `lock.\textsubscript{PL-GEN}'\\
    etc. \ldots

\rr{is this clear that the paradigm continues, but we don't show it? One of the reviewers thought the example was incomplete}

@@

\subsection{Corpus}
In order to evaluate the grammar we hand-annotated 10,150 words of Russian text
from Wikipedia articles, public domain literature and freely-available news sources. The
annotated texts are available online under the {\sc cc-by-sa} licence.\footnote{\url{https://svn.code.sf.net/p/apertium/svn/languages/apertium-rus/texts/}}\rr{is this a good url? Do we need to put a COPYING file in that folder?}

@@

\subsection{Qualitative evaluation}
%FRAN

In this section, we give a qualitative evaluation of errors made by the CG.

@@

\subsection{Quantitative evaluation}
% ROB accent placement task
